---
params: 
  title: ""
  publication_date: ""
  doi: ""
output: 
  html_document:
    anchor_sections: false
    theme: null
    highlight: null
    mathjax: null
    css: ["style.css", "https://fonts.googleapis.com/css?family=Source+Sans+Pro:400,700&display=swap"]
    self_contained: true
title: "`r params$title`"
---

```{r general-setup, include=FALSE}
## This file contains the ENGLISH version of the data story

# Set general chunk options
knitr::opts_chunk$set(echo = FALSE, fig.showtext = TRUE, fig.retina = 3, 
                      fig.align = "center", warning = FALSE, message = FALSE)

# Install pacman package if needed
if (!require("pacman")) {
  install.packages("pacman")
  library(pacman)
}

# Install snf.datastory package if not available, otherwise load it
if (!require("snf.datastory")) {
  if (!require("devtools")) {
    install.packages("devtools")
    library(devtools)
  }
  install_github("snsf-data/snf.datastory")
  library(snf.datastory)
}

# Load packages
p_load(tidyverse,
       lubridate,
       scales, 
       conflicted, 
       jsonlite,
       here, 
       ggiraph)

# Conflict preferences
conflict_prefer("filter", "dplyr")
conflict_prefer("get_datastory_theme", "snf.datastory")
conflict_prefer("get_datastory_scheme", "snf.datastory")

# Increase showtext package font resolution
showtext_opts(dpi = 320)

# Set the locale for date formatting (Windows)
Sys.setlocale("LC_TIME", "English")

# Create function to print number with local language-specific format 
print_num <- function(x) snf.datastory::print_num(x, lang = "de")

# Knitr hook for local formatting of printed numbers
knitr::knit_hooks$set(
  inline <- function(x) {
    if (!is.numeric(x)) {
      x
    } else {
      print_num(x)
    }
  }
)

# snf.plot::get_fancy_legend is not yet available in snf.datastory
get_fancy_legend <- function (direction = "horizontal", reverse = FALSE, ncol = NULL, 
  nrow = NULL, label_position = "right", title_position = "left", 
  label_hjust = 1)  {
  if (direction == "horizontal") {
    keywidth <- 8
    keyheight <- 2
  }
  else if (direction == "vertical") {
    keywidth <- 2
    keyheight <- 8
  }
  guide_legend(title.position = title_position, direction = direction, 
    keywidth = unit(keywidth, units = "mm"), keyheight = unit(keyheight, 
      units = "mm"), label.hjust = label_hjust, label.vjust = 0.2, 
    title.vjust = 0, byrow = FALSE, label.position = label_position, 
    reverse = reverse, ncol = ncol, nrow = nrow)
}
```

```{r print-header-infos, results='asis'}
# Add publication date to header
cat(format(as_datetime(params$publication_date), "%d.%m.%Y"))

# Create the CSV files if they don't exist
if (!file.exists(here("data", "current_panel_composition_20211109.csv")) |
    !file.exists(here("data", "hist_memberships_months_20211109.csv")))
  source(here("data-gathering.R"))

# Load the CSV files
current_memberships <- 
  read_csv(here("data", "current_panel_composition_20211109.csv"))
current_memberships_sum_stats <- 
  read_csv(here("data", "current_panel_composition_summarystats_20211109.csv"))
hist_memberships_months <- 
  read_csv(here("data", "hist_memberships_months_20211109.csv"))
hist_memberships_numbers_gender <- 
  read_csv(here("data", "hist_memberships_numbers_gender_20211109.csv"))
```

```{r story-specific-setup, include=FALSE}
# Set story-specific variables etc. here

## Calculate some figures to be used in the text

# Get the number of active distinct evaluation panels
number_of_panels <- current_memberships %>% 
  distinct(EvaluationBodyId_anon) %>% 
  nrow()

# Unique members
distinct_panelmembers <- current_memberships %>% 
  distinct(PersonId_anon) %>% 
  nrow() # 1'019 distinct persons

# Members in multiple panels
distinct_panelmembers_in_multiple_panels <- current_memberships %>%  
  group_by(PersonId_anon) %>%  
  summarise(multiple_panels = (n_distinct(EvaluationBodyId_anon) > 1)) %>% 
  count(multiple_panels) %>% 
  mutate(share = 100 * (n / sum(n))) %>% 
  filter(multiple_panels == TRUE) %>% 
  pull(share)

# Average age (overall)
mean_age_overall <- current_memberships_sum_stats$mean_age_overall

# Average age (male)
mean_age_male <- current_memberships_sum_stats$mean_age_male

# Average age (female)
mean_age_female <- current_memberships_sum_stats$mean_age_female

# Female share at reporting day (unique persons)
female_share_persons_reportingday <-
  current_memberships_sum_stats$female_share_persons_reportingday

# Female share at reporting day (memberships)
female_share_panels_reportingday <- 
  current_memberships_sum_stats$female_share_panels_reportingday

# # Gender shares
# distinct_panelmembers_gender <- current_memberships %>%
#   distinct(PersonId_anon, Gender) %>%
#   count(Gender) %>%
#   mutate(perc = 100 * (n / sum(n)))
# 
# # Share of external members (not memberships)
# share_external_members <- current_memberships %>%
#   group_by(PersonId_anon) %>%
#   # When this person has at least one role that is not "external member", it is
#   # not an external member.
#   summarise(is_internal = any(Function_En != "external member")) %>%
#   ungroup() %>%
#   count(is_internal) %>%
#   mutate(perc = 100 * (n / sum(n)))
```

__The SNSF's National Research Council decides whether or not to fund applications. The `r prettyNum(number_of_panels, big.mark = "'")` evaluation panels handle the preparatory work on which it bases its decisions, assessing several thousand applications each year.__

One of the key tasks of the evaluation panels is to ease the Research Council’s workload. Around 2010, the SNSF experienced a surge in applications. This in turn created a problem, as the 100 honorary experts who make up the committee had to increase their time commitment. Although some funding schemes already had evaluation panels, this was not enough. The situation prompted the SNSF to create other such panels, comprising members of the Research Council and university researchers.

The evaluation panels not only lighten the Research Council’s workload, they create added value by enhancing existing expertise and making it easier to adapt it to changing requirements. In addition, involving other researchers makes it easier to avoid conflicts of interest.

As of mid-November 2021, the `r prettyNum(number_of_panels, big.mark = "'")` panels had `r prettyNum(distinct_panelmembers, big.mark = "'")` members. `r  round(current_memberships_sum_stats$distinct_panelmembers_gender_m_perc)` per cent of whom are men and `r round(current_memberships_sum_stats$distinct_panelmembers_gender_f_perc)` per cent are women. `r round(current_memberships_sum_stats$share_external_members_external_perc)` per cent are from outside the Research Council. The average age is `r mean_age_overall` (`r mean_age_male` for men and `r mean_age_female` for women). `r round(distinct_panelmembers_in_multiple_panels)` per cent are members of several evaluation bodies.

<div class="plot-box">

<div class="plot-title">Active members and memberships</div>

```{r plot-member-development, out.width="100%", fig.height=4}
## Plot the development of # memberships and # members
# Summarise distinct persons and memberships data on a monthly level
memberships_numbers <- hist_memberships_months %>% 
  group_by(date) %>% 
  summarize(n_unique_members = n_distinct(PersonId_anon),
            n_memberships = n_distinct(Person_EvaluationBodyId_anon))

# Create the plot
p_memb <- memberships_numbers %>% 
  pivot_longer(c("n_unique_members", "n_memberships")) %>% 
  mutate(
    name = ifelse(name == "n_unique_members", "Individual members", 
                  "Memberships"),
    name = fct_relevel(name, c("Memberships", "Individual members"))
  ) %>% 
  filter(date >= ymd("2012-01-01"), 
         date <= ymd("2021-06-30")) %>% 
  ggplot(aes(x = date, y = value, color = name)) +
  labs(x = NULL,  y = NULL) +
  geom_line() +
  # Hack to add a (nearly) invisible hover point for every data point, in 
  # order to define a tooltip with ggiraph. Tooltips for geom_line_interactive
  # support only one distinct tooltip for the whole line, therefore, we use
  # geom_point here. 
  geom_point_interactive(
    aes(tooltip = paste0("<b>", format(date, "%B %Y"), "</b>\n", 
                         prettyNum(value, big.mark = "'"), " ", 
                         name)), size = 20, 
    color = "white", 
    alpha = 0.01) +
  scale_color_manual(values = get_datastory_scheme(), 
                     guide = get_fancy_legend()) +
  scale_x_date(
    limits = c(ymd("2012-01-01"), ymd("2021-06-30")),
    date_breaks = "years", date_labels = "%Y") +
  scale_y_continuous(limits = c(0, 1350), breaks = seq(0, 2000, 250)) + 
  get_datastory_theme()

# Create ggiraph object
girafe(ggobj = p_memb, 
       height_svg = 4, 
       options = list(
         opts_toolbar(saveaspng = FALSE),
         opts_hover(css = "fill:#F75858;stroke:#F75858;"),
         opts_tooltip(
           css = get_ggiraph_tooltip_css(),
           opacity = 0.8,
           delay_mouseover = 0,
           delay_mouseout = 0
         )
       ))
```

<div class="caption">
One person may be a member of several evaluation panels. <br/>
Source: SNSF internal database, excerpt available <a href="https://github.com/snsf-data/datastory_evaluation_panels/tree/main/data" target="_blank">here</a>. Past values are not definite, see “Data and methods” box.
</div>

<div class="info-box">

### SNSF evaluation procedure

The SNSF uses a peer review procedure to evaluate the projects submitted by researchers. In most funding schemes, applications that satisfy the formal criteria are first sent to suitably qualified external reviewers, many of whom are in a different country. This is followed by an internal evaluation, during which a referee evaluates the application using the external reviewer’s assessment and submits it to the responsible evaluation panels with a recommendation. The panel assesses the application on the basis of the scientific criteria for the scheme in question and ranks it against other applications. The responsible superordinate committee of the Research Council then makes a provisional decision on which projects to fund. The final decision is made by the Presiding Board of the Research Council.

</div>

```{r ch-or-aborad-shares-2012-2021}
# Get the share of CH/abroad members of today
ch_or_abroad_2021 <- current_memberships %>% 
  filter(!is.na(Country), Country != "Unknown") %>% 
  distinct(PersonId_anon, Country) %>%
  mutate(abroad = Country != "Switzerland") %>% 
  count(abroad) %>% 
  mutate(perc = 100 * (n / sum(n)))

# Get the share of CH/abroad members of January 2012
ch_or_abroad_2012 <- hist_memberships_months %>% 
  filter(!is.na(Country), Country != "Unknown", 
         date == "2012-01-01") %>% 
  distinct(PersonId_anon, Country) %>%
  mutate(abroad = Country != "Switzerland") %>% 
  count(abroad) %>% 
  mutate(perc = 100 * (n / sum(n)))
```


### Greater internationality

To expand the evaluation panels, the SNSF invited an increasing number of researchers from outside Switzerland to contribute. In addition to easing workload, this also has the advantage of further reducing conflicts of interest. Today, `r ch_or_abroad_2021 %>% filter(abroad) %>% pull(n)` – or `r ch_or_abroad_2021 %>% filter(abroad) %>% pull(perc) %>% round()` per cent – of panel members work abroad. At the start of 2012, the equivalent figure was `r ch_or_abroad_2012 %>% filter(abroad) %>% pull(perc) %>% round()` per cent or just `r ch_or_abroad_2012 %>% filter(abroad) %>% pull(n)` members.

Country representation has also evolved over time. While the percentage of US Americans has almost halved, there are significantly more members from Italy, Great Britain and the Netherlands in percentage terms than was the case just under 10 years ago.

<div class="plot-box">

<div class="plot-title">International composition of bodies, in per cent</div>

```{r plot-countryshare-development, out.width="100%", fig.height=5}
# Create a mapping table of countries and their "group" (country name or 
# "Other") to be published in the plot. Put countries with overall small 
# occurrence into "Other".
country_groups <- hist_memberships_months %>% 
  filter(!is.na(Country), Country != "Unknown") %>% 
  distinct(PersonId_anon, Country) %>% 
  count(Country) %>% 
  arrange(-n) %>% 
  mutate(freq = n / sum(n), 
         # Under 1% -> Other
         country_group = ifelse(freq < 0.01, "Other countries", Country)) %>% 
  select(-freq, -n)

# Prepare the data for the area plot
memberships_numbers_country <- hist_memberships_months %>% 
  # Remove CH
  filter(Country != "Switzerland") %>% 
  # Get the country group, resp. the "Other"
  left_join(country_groups, by = "Country") %>% 
  # Remove entries where we don't know the country
  filter(!is.na(country_group))

# Now sumarize the data for the plot
memberships_numbers_country <- memberships_numbers_country %>% 
  group_by(date, country_group) %>% 
  # Persons, not memberships
  summarise(n_memberships = n(), 
            n_unique_members = n_distinct(PersonId_anon)) %>% 
  mutate(freq_unique = n_unique_members / sum(n_unique_members), 
         date_year = year(date)) %>% 
  ungroup() %>% 
  # For the tooltip: Create one tooltip per country and year (and not per 
  # country, year and month)
  left_join(
    memberships_numbers_country %>% 
      mutate(date_year = year(date)) %>% 
      group_by(date_year, country_group) %>% 
      summarise(year_unique_members = n_distinct(PersonId_anon)) %>% 
      mutate(freq_unique_year = year_unique_members / 
               sum(year_unique_members)) %>% 
      ungroup(), by = c("country_group" = "country_group", 
                        "date_year" = "date_year"))

# Color scheme for the plot
colors <- c("#5298BD",  "#FBBE5E", "#71B294", "#9D90B9", "#CBE0EB", "#B2B1A7", 
            "#FDDEAE", "#83D0F5", "#B8D8C9", "#CEC7DC", "#D8D8D3", "#DAF1FC")

# Create the area plot
p_country_shares <- memberships_numbers_country %>% 
  ggplot(aes(x = date, y = freq_unique, fill = country_group)) + 
  scale_x_date(limits = c(ymd("2012-01-01"), ymd("2022-01-01")), 
               breaks = seq(ymd("2012-01-01"), ymd("2022-01-01"), "year"), 
               date_labels = "%Y") + 
  geom_area() + 
  # Hack: geom_area_interactive() has visual corruption when providing dynamic
  # tooltip, draw the corrupted area plot with nearly no visibility over the 
  # visible geom_area() underneath.
  geom_area_interactive(aes(tooltip = 
                              paste0("<b>", country_group, 
                                     "</b> (", date_year,
                                     ")<br>",
                                     year_unique_members , " ",
                                     " members located (",
                                     round(freq_unique_year * 100),
                                     "%)<br>in ", 
                                     ifelse(country_group == "Other countries",
                                            "other countries", country_group), 
                                     ".")), 
                        alpha = 0.01) +
  scale_fill_manual(values = colors, 
                    guide = get_fancy_legend(direction = "vertical", ncol = 1, 
                                             label_hjust = 0)) +
  get_datastory_theme(legend_position = "right") + 
  scale_y_continuous(labels = percent_format(accuracy = 1)) +
  labs(x = NULL, y = NULL) 

# Create ggiraph object
girafe(ggobj = p_country_shares,
       height_svg = 5,
       options = list(
         opts_toolbar(saveaspng = FALSE),
         opts_hover(css = "fill:#F75858;stroke:#F75858;"),
         opts_tooltip(
           css = get_ggiraph_tooltip_css(),
           opacity = 0.8,
           delay_mouseover = 0,
           delay_mouseout = 0
         )
       ))
```

<div class="caption">
Source: SNSF internal database, excerpt available <a href="https://github.com/snsf-data/datastory_evaluation_panels/tree/main/data" target="_blank">here</a>. Past values are not definite, see “Data and methods” box.
</div>
</div>

### Percentage of women rising too slowly – quota introduced

Although the number of women in the evaluation panels has constantly risen, at `r round(female_share_persons_reportingday)` per cent it is still significantly below the figure for men. Counting memberships rather than people produces the same picture, with women holding `r round(female_share_panels_reportingday)` per cent of memberships. This is a reflection of their under-representation in top scientific positions in Switzerland. <a href="https://www.snf.ch/de/kjCKYzJgvuNWbsf2/news/news-210205-quoten-fuer-mehr-gleichstellung-in-der-forschung" target="_blank">While the number of women is still just under 45 per cent at PhD level, the figure falls continuously with progressing career levels; just 23 per cent of Swiss professorships are held by women</a>.

<div class="plot-box">

<div class="plot-title">Panel members by gender</div>

```{r plot-gendershare-development, out.width="100%", fig.height=3}
# Create the lineplot with the gender shares
p_memberships_numbers_gender <- hist_memberships_numbers_gender %>%
  mutate(Gender = ifelse(Gender == "f", "Women", "Men")) %>% 
  ggplot(aes(x = date, y = freq_unique, color = Gender)) +
  geom_line() + 
  # Hack to add a (nearly) invisible hover point for every data point, in 
  # order to define a tooltip with ggiraph. Tooltips for geom_line_interactive
  # support only one distinct tooltip for the whole line, therefore, we use
  # geom_point here. 
  geom_point_interactive(
    aes(tooltip = paste0("<b>", format(date, "%B %Y"), "</b>\n", 
                         prettyNum(round(freq_unique * 100), big.mark = "'"), 
                         "% ", Gender, " (", n_unique_members, " members)")), 
    size = 20, color = "white", alpha = 0.01) +
  scale_x_date(limits = c(ymd("2012-01-01"), "2021-08-31"), 
               date_breaks = "years", date_labels = "%Y")  + 
  scale_color_manual(values = get_datastory_scheme()) + 
  labs(x = NULL, y = NULL) + 
  scale_y_continuous(labels = percent_format(accuracy = 1), limits = c(0, 1)) +
  get_datastory_theme()

# Create ggiraph object
girafe(ggobj = p_memberships_numbers_gender,
       height_svg = 4,
       options = list(
         opts_toolbar(saveaspng = FALSE),
         opts_hover(css = "fill:#F75858;stroke:#F75858;"),
         opts_tooltip(
           css = get_ggiraph_tooltip_css(),
           opacity = 0.8,
           delay_mouseover = 0,
           delay_mouseout = 0
         )
       ))
```

<div class="caption">
Source: SNSF internal database, excerpt available <a href="https://github.com/snsf-data/datastory_evaluation_panels/tree/main/data" target="_blank">here</a>. Past values are not definite, see “Data and methods” box.
</div>
</div>

To redress this imbalance, the Presiding Board of the Research Council decided to adopt a gender quota for its evaluation panels. This must have been achieved no later than the end of the transition period in 2026. For subject-specific panels, the quota takes into account the proportion of women in the respective subject.


<div class="info-box">

### Data and methods

The data used in this article comes from the SNSF's internal database. The data can be obtained <a href="https://github.com/snsf-data/datastory_evaluation_panels/tree/main/data" target="_blank">from Github</a>.

* Data quality issues were encountered when determining figures for the past. In some cases, there is no clear information on when a particular person joined or left a panel. Where this has occurred, we have adopted the average membership period. In a small number of cases, datasets did not give either the start or end of membership. These cases have been excluded.

* The country assigned to panel members is the country of the institute of which they were/are members. Where this was not known, we relied on the country code in their e-mail address.

* Members of the former SNSF Research Commissions at universities have not been included.

* Analysis carried out on 9 November 2021.



</div>

Details of current panel members can be found on the <a href="https://www.snf.ch/de/Kh7DZxlDCAXUljIZ/seite/ueberuns/organisation/forschungsrat-und-evaluationsgremien/mitglieder-der-evaluationsgremien" target="_blank">Members of the evaluation bodies page</a> of our website.

Data, text and code of this data story are <a href="https://github.com/snsf-data/datastory_evaluation_panels" target="_blank">available on Github</a> and <a href="https://doi.org/10.46446/datastory.central-role-of-evaluation-panels" target="_blank">archived on Zenodo</a>.